---
output:
  html_document: default
  pdf_document: default
---

<!-- # Filename: Eugene_Police_Budget_Report.rmd -->

<!-- # Author: @russl_corey -->

<!-- # Date: oct 10, 2022 -->

<!-- # This program is free software: you can redistribute it and/or modify it under  -->

<!-- # the terms of the GNU General Public License as published by the Free Software  -->

<!-- # Foundation, either version 3 of the License, or (at your option) any later  -->

<!-- # version. -->

<!-- #  -->

<!-- # This program is distributed in the hope that it will be useful, but WITHOUT ANY  -->

<!-- # WARRANTY; without even the implied warranty of MERCHANTABILITY or FITNESS FOR A  -->

<!-- # PARTICULAR PURPOSE. See the GNU General Public License for more details. -->

<!-- # You should have received a copy of the GNU General Public License along with  -->

<!-- # this program. If not, see <https://www.gnu.org/licenses/>.  -->

---
title: "Eugene, Oregon Police Budget"
author: Russell_Corey
date: December, 8 2022
output: html_document
---

\newpage

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(ggplot2)
library(readr)
library(scales)
library(dbplyr)
```

```{r Load Data, echo=FALSE, message=FALSE, warning=FALSE}
setwd('/home/russell/Dropbox/DataAnalysis/Eugene_Police')
source("police_budget.r")
```

```{r Load Offenses Data, echo=FALSE, message=FALSE, warning=FALSE}
setwd('/home/russell/Dropbox/DataAnalysis/Eugene_Police')
source('OpenData-Offenses-All.r')
```
# Intro

 Source code will be posted on [Github](https://github.com/russl-corey/Oregon_Crime)

# Data

"Open Budget Eugene"

    https://budget.eugene-or.gov/#!/year/2018/operating/0/department/Police/0/service?vis=barChart

\newpage

# Plots



```{r Plot Police Budget, echo=FALSE, message=FALSE, results='asis'}



ggplot() +
  labs(y='Dollars', title="Eugene Police Adopted Budget") +
  scale_x_date(date_labels = "%Y", date_breaks = "1 year") +
  geom_line(data=budget[budget$Department == 'Police',], 
            aes(x=year, y=`Adopted Budget`, color='Police'))+
  scale_y_continuous(labels = comma)
  # geom_line(data=budget_all[budget_all$Department == 'Public Works',], 
  #           aes(x=year, y=`Adopted Budget`, color='Public Works'))+
  # geom_line(data=budget_all[budget_all$Department == 'Fire and Emergency Medical Services',], 
  #           aes(x=year, y=`Adopted Budget`, color='Fire and Emergency Medical Services'))+
  # geom_line(data=budget_all[budget_all$Department == 'Library  Recreation and Cultural Services',], 
  #           aes(x=year, y=`Adopted Budget`, color='Library  Recreation and Cultural Services'))

budget %>%
  filter(Department == 'Police') %>%
  dplyr::select(c('year', 'Adopted Budget')) %>%
  kable(caption = 'Police Budget', format.args = list(big.mark = ",",
  scientific = FALSE))

budget_police %>%
  filter(Service == 'Patrol') %>%
  dplyr::select(c('year', 'Adopted Budget')) %>%
  kable(caption = 'Police Patrol Budget', format.args = list(big.mark = ",",
  scientific = FALSE))

```



```{r Plot Police Budget Increase, echo=FALSE, results = 'asis', warning=FALSE, }

data <- budget %>%
  filter(Department == 'Police') %>%
  dplyr::select('year', 'budget_increase') 

  barplot(data$budget_increase,
          main="Budget increase over previous year",
          names.arg = data$year,
          ylab='dollars',
          las =2,
          cex.names = 0.7,
          format.args = list(big.mark = ",",
  scientific = FALSE)
)
#   
# data <- budget %>%
#   filter(Department == 'Police') %>%
#   dplyr::select('year', 'budget_increase_percnt') 
# 
#   barplot(data$budget_increase,
#           main="Budget increase over previous year",
#           names.arg = data$year,
#           ylab='% increase',
#           las =2,
#           cex.names = 0.7
# )

```


```{r Plot Eugene Arrests, echo=FALSE, message=FALSE, results='asis'}

violent_titles <- c("Aggravated Assault",
                    "Forcible Rape", 
                    "Forcible Sodomy", 
                    "Robbery",
                    'Willful Murder')
# Set the working directory to where the data sets are located
folder <- "~/Documents/Oregon_Crime_Report/"

# Load data
arrests <- read_csv("~/Documents/Eugene_Police/data/OpenData-Arrests-All.csv")

# Format data
arrests$agency <- tolower(arrests$`Agency Name`)
arrests$date <- as.Date(arrests$Date, format = "%m/%d/%Y", origin="1900-01-01")


arrests[grep('eugene', arrests$agency),] %>%
  group_by(month=floor_date(date, 'month')) %>%
  summarize(arrests = sum(`Distinct Arrest Offenses`)) %>%
  ggplot() +
  labs(y='arrests', title="Eugene Police Arrests") +
  scale_x_date(date_labels = "%Y", date_breaks = "6 month") +
  scale_y_continuous(labels = comma)+
  geom_line(aes(x=month, y=arrests, color='Total Arrests'))
  

```

```{r Plot Eugene Offenses, echo=FALSE, message=FALSE}

eugene_o <- offenses[grep('eugene', offenses$agency), ]

o_month <- eugene_o %>%
  group_by(month=floor_date(date, 'month')) %>%
  summarize(o_sum = sum(`Distinct Offenses`))

vo_month <- eugene_o %>%
  filter(vo == 1) %>%
  group_by(month=floor_date(date, 'month')) %>%
  summarize(vo_sum = sum(`Distinct Offenses`))

ggplot() +
  labs(y='Dollars', title="Eugene Distinct Offenses") +
  scale_x_date(date_labels = "%Y", date_breaks = "6 month") +
  geom_line(data=o_month, aes(x=month, y=o_sum, color='All Offenses'))+
  geom_line(data=vo_month, aes(x=month, y=vo_sum, color='Violent Offenses'))+
  scale_y_continuous(labels = comma)

eugene_o  %>%
  head() %>%
  kable(caption = 'Police Budget', format.args = list(big.mark = ",",
  scientific = FALSE))

```


\newpage

# Sources

#### Eugene Police Budget

"Open Budget Eugene"

    https://budget.eugene-or.gov/#!/year/2018/operating/0/department/Police/0/service?vis=barChart






\newpage

# Appendix B

source code available at <https://github.com/russl-corey/Oregon_Crime>

This program is free software: you can redistribute it and/or modify it under the terms of the GNU General Public License as published by the Free Software Foundation, either version 3 of the License, or (at your option) any later version.

This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE. See the GNU General Public License for more details. You should have received a copy of the GNU General Public License along with this program. If not, see [\<https://www.gnu.org/licenses/\>](https://www.gnu.org/licenses/).
